---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r loading}
library(lubridate)
library(datasets)
library(dplyr)

data <- read.csv(unz("activity.zip","activity.csv",open = "", encoding = getOption("encoding")), na.strings ="NA")
data$date <- ymd(data$date)
data$empty <- is.na(data$steps)
data_clean <- subset(data, !empty)

summary(data)
```
## What is mean total number of steps taken per day?

#Create aggregated steps per day  
```{r aggregatedays, results = "asis"}
daily_steps <- data_clean %>% group_by(date) %>% summarize(daily_steps_sum = sum(steps))


# Create histogram  
hist(
  daily_steps$daily_steps_sum, 
  xlab = "daily steps taken", 
  main = "Histogram Total Steps each Day", 
  col = "blue")
mean_steps <- mean(daily_steps$daily_steps_sum)
median_steps <- median(daily_steps$daily_steps_sum)

```
`r mean_steps` is the mean & `r median_steps` is the median on steps taken each day

## What is the average daily activity pattern?
```{r aggregateinterval}
interval_avg <- data_clean %>% group_by(interval) %>% summarize(avg_steps = mean(steps))
with(
  interval_avg, 
  plot(
    interval, 
    avg_steps, 
    type = "l", 
    main = "Average 5 Min steps over all days", 
    xlab = "Interval, hhmm"))

max_step_interval <- interval_avg$interval[which.max(interval_avg$avg_steps)]

```
Over all days, `r max_step_interval` is the interval containing the maximum number of steps  


## Imputing missing values
```{r missing}
missing_value <- sum(is.na(data$steps))

```
There are `r missing_value` NA found in rows of the data set.  
  
To fill in the NAs, we take the average number of steps during that 5 minute interval over all days.
```{r replaceNA}
fill_data <- data
for (i in 1:nrow(fill_data)) {
    if (is.na(fill_data$steps[i])) {
        # Find index value for interval = average
        myindex <- which(fill_data$interval[i] == interval_avg$interval)
        # replace the NA
        fill_data$steps[i] <- interval_avg[myindex,]$avg_steps
    }
}
fill_data$date <- ymd(fill_data$date)

filldailysteps <-  fill_data %>% group_by(date) %>% summarize(filldaily_steps_sum = sum(steps))

# Create histogram  
hist(
  filldailysteps$filldaily_steps_sum, 
  xlab = "daily steps taken", 
  main = "Histogram Total Steps each Day", 
  col = "purple")

fillmean_steps <- mean(filldailysteps$filldaily_steps_sum)
fillmedian_steps <- median(filldailysteps$filldaily_steps_sum)
```
`r fillmean_steps` is the mean & `r fillmedian_steps` is the median on steps taken each day with filled NA rows.  
The median is now equal to mean.  


## Are there differences in activity patterns between weekdays and weekends?

```{r}
fill_data$weekday <- weekdays(fill_data$date)
fill_data$daytype <- "weekday"
fill_data$daytype[fill_data$weekday %in% c("Samstag", "Sonntag")] <- "weekend"

dayaverage <- fill_data %>%
    group_by(daytype, interval) %>%
    summarize(avg_steps=mean(steps))


library(ggplot2)

g <- ggplot(dayaverage, mapping = aes(interval, avg_steps))
g + geom_line() + facet_grid(rows = vars(daytype ))+labs(x = "Interval", y = "Number of Steps") + theme_light()

```
